---
title: "Final_Activity_v3"
author: "Foster Thorburn"
date: "2022-11-30"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
#edit test
library(ggplot2)
library(dplyr)
library(RWDataPlyr)
library(readxl)
library(skimr)
library(forcats)
library(janitor)
library(stringr)
library(tidyr)
library(ggthemes)
```

## Data Dictionary (after recoding)
```{r}

variable <- c("FIPS_code", "GTCSA", "Stabr", "Age", "Race", "Sex", 
                     "Marital_status", "High_Lvl_Degree", "twelveM_Household_FI", "twelveM_Household_FI_Rasche", "FI_orNot")

description <- c("Unique County Code", "Consolidated Statistical Area (CSA) FIPS code", "Abreviated State Name", "Age", "Race", "Sex", "Marital Status","Highest Level Ed. Degree Completeed", "Household Food Security Scale, 12-Month Reference Period", "Food Security Rasch Scale Score, 
12-Month Recall", "High Food Insecurity or not")

type <- c("Categorical", "Categorical", "Categorical", "Quantitative", "Categorical", "Categorical", "Categorical", "Categorical", "Quantitative", "Categorical", "Categrorical")

data_dec21_dictionary <- data.frame(variable, description, type)

data_dec21_dictionary
  
```



## Reading in Current Population Survey, December 2021: Food Security Supplement
```{r data}

# Read in csv file
dec21pub <- read.csv("C:/Users/foste/Desktop/418 final project folder/dec21pub.csv")

```

Subsetting Food Supplement data set for variables          
```{r 2}



data_dec21 <- select(dec21pub, GTCO, GTCSA, PRTAGE, PTDTRACE, PESEX, 
                     PEMARITL, GESTFIPS, PEEDUCA,PEHSPNON, HRFS12M4, HRFS12M1, HRFS12MD)

#removing old F.S data sets
rm(dec21pub)

```


Creating unique County FIPS codes
```{r}
                            #state    #county
# Unique County FIPS code = [GESTFIPS][GTCO]

#SATE-CODE

#test var holds the value of the length of the state code
data_dec21 <- data_dec21 %>% 
          mutate(test_var = nchar(as.character(factor(GESTFIPS))))

#code to add one zero to the front of the state code if there is only one value, if there is two adds nothing
data_dec21 <- data_dec21 %>% 
  mutate(test_var2 = ifelse(data_dec21$test_var == 1, str_c("0",data_dec21$GESTFIPS,sep=""),data_dec21$GESTFIPS))


#COUNTY-CODE

#test_var_county holds the value of the length of the county code
data_dec21 <- data_dec21 %>% 
          mutate(test_var_county = nchar(as.character(factor(GTCO))))


#code to add one zero to the front of the county code if there is less than or equal to two values 
  data_dec21 <- data_dec21 %>% 
  mutate(test_var2_county = ifelse(data_dec21$test_var_county <= 2, str_c("0",data_dec21$GTCO,sep=""),data_dec21$GTCO))

#Reruning the counter variable to find the length of the updated county code
data_dec21 <- data_dec21 %>% 
          mutate(test_var_county = nchar(test_var2_county))

#Iterating a second time to add another zero if there's only two values, otherwise leave value be  
data_dec21 <- data_dec21 %>% 
  mutate(test_var2_county = ifelse(data_dec21$test_var_county <= 2, str_c("0",data_dec21$test_var2_county,sep=""),data_dec21$test_var2_county))

#UNIQUE-COUNTY CODE

data_dec21 <- data_dec21 %>% 
          mutate(FIPS_code = str_c(test_var2,test_var2_county,sep=""))



#example function
#my_funct <- function() {
#  print("test")
#}

```


Re-coding Demographic Variables
```{r} 


#Replacing -1 values with NA for entire data set
data_dec21 <- data_dec21 %>%
                  mutate(across(everything(), ~na_if(str_trim(.x), "-1")))
#Replacing -9 values with NA for entire data set
data_dec21 <- data_dec21 %>%
                  mutate(across(everything(), ~na_if(str_trim(.x), "-9")))
#Replacing -6 values with NA for entire data set
data_dec21 <- data_dec21 %>%
                  mutate(across(everything(), ~na_if(str_trim(.x), "-6")))

#Recoding Hispanic Variable - PEHSPNON
data_dec21 <- data_dec21 %>% 
  mutate(Hispanic = factor(PEHSPNON) %>% 
    fct_recode(
      "Hispanic" = "1",
      "Non-Hispanic" = "2"))


#Recoding CSA Area variable
data_dec21 <- data_dec21 %>% 
  mutate(CSA_Area = factor(GTCSA) %>% 
    fct_recode(
      "Grand Rapids-Wyoming" = "24340",
      "Detroit-Warren-Dearborn" = "24340",
      "Ann Arbor" = "11460",
      "Lansing-East Lansing" = "29620"))

#Re-coding marital status variable
data_dec21 <- data_dec21 %>% 
  mutate(Marital_status = factor(PEMARITL) %>% 
    fct_recode(
      "MARRIED - SPOUSE PRESENT" = "1",
      "MARRIED - SPOUSE ABSENT" = "2",
      "WIDOWED" = "3",
      "DIVORCED" = "4",
      "SEPARATED" = "5",
      "NEVER MARRIED" = "6"))

#Re-coding highest complete degree variable
data_dec21 <- data_dec21 %>% 
  mutate(High_Lvl_Degree = factor(PEEDUCA) %>% 
    fct_recode(
      "Less than H.S or GED" = "31",
      "Less than H.S or GED" = "32",
      "Less than H.S or GED" = "33",
      "Less than H.S or GED" = "34",
      "Less than H.S or GED" = "35",
      "Less than H.S or GED" = "36",
      "Less than H.S or GED" = "37",
      "Less than H.S or GED" = "31",
      "Less than H.S or GED" = "38",
      "H.S or GED" = "39", 
      "H.S or GED" = "40",
      "Associate's" = "41",
      "Associate's" = "42",
      "Bachelor's" = "43",
      "Master's" = "44",
      "Professional" = "45",
      "Doctorate" = "46"))


#Makes state variable GESTFIPS into character
data_dec21$GESTFIPS <- as.character(data_dec21$GESTFIPS)
#summary(data_dec21$GESTFIPS)

#Changing data set to make a new column called "Stabr" which is the state abreviation
data_dec21 <- data_dec21 %>%
                   mutate(Stabr = factor(GESTFIPS) %>%
                     fct_recode( 
           "AL" = "1", 
           "AK" = "2", 
           "AZ" = "4",
           "AR" = "5",
           "CA" = "6",
           "CO" = "8",
           "CT" = "9",
           "DE" = "10",
           "DC" = "11",
           "FL" = "12",
           "GA" = "13",
           "HI" = "15",
           "ID" = "16",
           "IL" = "17",
           "IN" = "18",
           "IA" = "19",
           "KS" = "20",
           "KY" = "21",
           "LA" = "22",
           "ME" = "23",
           "MD" = "24",
           "MA" = "25",
           "MI" = "26",
           "MN" = "27",
           "MS" = "28",
           "MO" = "29",
           "MT" = "30",
           "ME" = "31",
           "MV" = "32",
           "NH" = "33",
           "NJ" = "34",
           "NM" = "35",
           "NY" = "36",
           "NC" = "37",
           "ND" = "38",
           "OH" = "39",
           "OK" = "40",
           "OR" = "41",
           "PA" = "42",
           "RI" = "44",
           "SC" = "45",
           "SD" = "46",
           "TN" = "47",
           "TX" = "48",
           "UT" = "49",
           "VT" = "50",
           "VA" = "51",
           "WA" = "53",
           "WV" = "54",
           "WI" = "55",
           "WY" = "56"))

#data_dec21 <- data_dec21 %>%
#                   mutate(new_name = factor(oldname) %>%
#                     fct_recode( 
#           "new_level" = "old_level")



#Recoding Race Variable 

data_dec21$PTDTRACE <- as.character(data_dec21$PTDTRACE)

data_dec21 <- data_dec21 %>% 
  mutate(Race = factor(PTDTRACE) %>% 
    fct_recode(
      "White Only" = "1",
      "Black Only" = "2",
      "American Indian, Alaskan Native Only" = "3",
      "Aisan Only" = "4",
      "Hawaiian/Pacific Islander Only" = "5",
      "White-Black" = "6",
      "White-AI" = "7",
      "White-Asian" = "8", 
      "White-HP" = "9",
      "Black-AI" = "10",
      "Black-Asian" = "11",
      "Black-HP" = "12",
      "AI-Asian" = "13",
      "AI-HP" = "14",
      "Asian-HP" = "15",
      "3+ mixed-race" = "16",
      "3+ mixed-race" = "17", 
      "3+ mixed-race" = "18", 
      "3+ mixed-race" = "19", 
      "3+ mixed-race" = "20", 
      "3+ mixed-race" = "21", 
      "3+ mixed-race" = "22", 
      "3+ mixed-race" = "23", 
      "3+ mixed-race" = "24", 
      "3+ mixed-race" = "25", 
      "3+ mixed-race" = "26"))



#Mutating to re-code Sex Variable 
data_dec21 <- data_dec21 %>% 
  mutate(Sex = factor(PESEX) %>% 
    fct_recode(
      "Male" = "1",
      "Female" = "2"))     

#Mutating to recode Age variable
 data_dec21 <- data_dec21 %>% 
  mutate(Age = as.numeric(factor(PRTAGE)))
 

                  
```
Re-coding Food Security Measures
```{r}
#Mutating to re code Main food insecurity measure variable
 data_dec21 <- data_dec21 %>% 
  mutate(twelveM_Household_FI = factor(HRFS12M1) %>% 
           fct_recode(
             "Food Secure High or Marginal Food Security" = "1",
             "Low Food Security" = "2",
             "Very Low Food Security" = "3"))
 
 #Mutating to recode food insecurity scale Rasche variable
 data_dec21 <- data_dec21 %>% 
  mutate(twelveM_Household_FI_Rasche = factor(HRFS12M4))
 #Making the Rasche variable numeric so it can be graphed
 
 data_dec21$twelveM_Household_FI_Rasche <- as.numeric(as.character(data_dec21$twelveM_Household_FI_Rasche))
 
 
 #Mutating to re code food insecurity variable which combines the main food insecurity measure into food insecure or not
 data_dec21 <- data_dec21 %>% 
  mutate(FI_orNot = factor(HRFS12MD) %>% 
           fct_recode(
             "High Food Security" = "1",
             "Food Insecure" = "2",
             "Food Insecure" = "3",
             "Food Insecure" = "4"))
```


Re-selecting Re-coded Variables for new data set

```{r}

data_dec21 <- select(data_dec21,FIPS_code, GTCSA, Stabr, Age, Race, Sex,Hispanic, 
                     Marital_status, High_Lvl_Degree, twelveM_Household_FI, twelveM_Household_FI_Rasche, FI_orNot)




```

Summary Statistics of characteristic variables

```{r}
#Age
summary(data_dec21$Age)

hist(data_dec21$Age, main = "Histogram of Age")

#Race

summary(data_dec21$Race)

data_dec21 %>%  filter(!is.na(Race)) %>% ggplot(aes(x = Race)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1)) + #45 degree angle labels 
  labs(title = "Histogram of Race Counts", x = "Race", caption = "Source: Current Population Survey Food Security Supplements")


#Sex
data_dec21 %>%  filter(!is.na(Sex)) %>% ggplot(aes(x = Sex)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1)) + #45 degree angle labels
  labs(title = "Histogram of Sex Counts", x = "Sex", caption = "Source: Current Population Survey Food Security Supplements")


#Marital Status
data_dec21 %>%  filter(!is.na(Marital_status)) %>% ggplot(aes(x = Marital_status)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1)) + #45 degree angle labels
  labs(title = "Histogram of Marital Status Counts", x = "Marital Status", caption = "Source: Current Population Survey Food Security Supplements")


#High Level Degree
data_dec21 %>%  filter(!is.na(High_Lvl_Degree)) %>% ggplot(aes(x = High_Lvl_Degree)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1)) + #45 degree angle labels
  labs(title = "Histogram of Highest Level Degree Counts", x = "Highest Level Degree", caption = "Source: Current Population Survey Food Security Supplements")

## Geographic

#State
data_dec21 %>%  filter(!is.na(Stabr)) %>% ggplot(aes(x = Stabr)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1)) + #45 degree angle labels
  labs(title = "Histogram of State Counts", x = "State", caption = "Source: Current Population Survey Food Security Supplements") +
  theme(axis.text.x = element_text(size = 7, face = "bold")) #making x-axis names bold and sliglty smaller for visibility

#Hispanic
data_dec21 %>%  filter(!is.na(Hispanic)) %>% ggplot(aes(x = Hispanic)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1)) + #45 degree angle labels
  labs(title = "Histogram of Hispanics", x = "Hispanic or not", caption = "Source: Current Population Survey Food Security Supplements") +
  theme(axis.text.x = element_text(size = 7, face = "bold")) #making x-axis names bold and sliglty smaller for visibility


```

Summary Statistics of Food Security Measures
```{r}

#Food Insecurity Main measure
data_dec21 %>%  filter(!is.na(twelveM_Household_FI)) %>% ggplot(aes(x = twelveM_Household_FI)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1)) + #45 degree angle labels
  labs(title = "Histogram of Main Food Insecurity Measure Counts", x = "Food Insecurity Level", caption = "Source: Current Population Survey Food Security Supplements")


#Food Insecurity
data_dec21 %>%  filter(!is.na(FI_orNot)) %>% ggplot(aes(x = FI_orNot)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1)) + #45 degree angle labels
  labs(title = "Histogram of High Food Insecurity or Not Counts", x = "Food Insecurity Level", caption = "Source: Current Population Survey Food Security Supplements")
```


Based on the histogram of Race I have decided to make some of the values mixed race rather than their own category due to their relatively extremely low numbers:

Other Mixed Race:
Black-Asian 
AI-HP
Black-HP
Asian-HP
White-HP
Black-AI
AI-Asian
3+ mixed race

White-Black 

White Asian

White-AI

White ONly 

Asian Only 

Black Only

American Indian, Alaskan Native Only 

Hawaiian/Pacific Islander Only

Re-coding the Race variable:
```{r}
data_dec21 <- data_dec21 %>% 
  mutate(Race = factor(Race) %>% 
    fct_recode(
      "Other Mixed Race" = "Black-Asian",
      "Other Mixed Race" = "AI-HP",
      "Other Mixed Race" = "Black-HP",
      "Other Mixed Race" = "Asian-HP",
      "Other Mixed Race" = "White-HP",
      "Other Mixed Race" = "White-HP",
      "Other Mixed Race" = "Black-AI",
      "Other Mixed Race" = "AI-Asian",
      "Other Mixed Race" = "3+ mixed-race"))

#Rerunning Race Histogram Code
data_dec21 %>%  filter(!is.na(Race)) %>% ggplot(aes(x = Race)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1)) + #45 degree angle labels 
  labs(title = "Histogram of Race Counts", x = "Race", caption = "Source: Current Population Survey Food Security Supplements")

summary(data_dec21$Race)

```

## Final Look at data_dec21 dataset
```{r}

skim(data_dec21)

glimpse(data_dec21)

```




## Looking at Food insecurity 


Rasche Scores
```{r, warning = FALSE}

#data_dec21$HRFS12M4 <- as.numeric(data_dec21$HRFS12M4)
#rasche_FIscore12mHoushold <-data_dec21$HRFS12M4
#data_dec21$twelveM_Household_FI_Rasche <- as.numeric(data_dec21$twelveM_Household_FI_Rasche)

#Graphing Food Insecurity Rasche scores by Age/Gender
data_dec21 %>% filter(!is.na(twelveM_Household_FI_Rasche)) %>% 
  ggplot(aes(x = Age, y = twelveM_Household_FI_Rasche, color = Sex)) +
  geom_jitter(alpha = .05) +
  stat_summary(aes(x = Age, y = twelveM_Household_FI_Rasche, group = Sex), fun.y=mean, geom="line", size =1.5) +
  labs(y = "Mean Food Insecurity Rasche Score (HRFS12M4)", title = "Mean Food Insecurity Rasche Scores by Age/Gender", caption = "Source: Current Population Survey Food Security Supplements") +
  theme_bw() 


data_dec21 %>% filter(!is.na(twelveM_Household_FI_Rasche)) %>% 
tabyl(Sex, twelveM_Household_FI_Rasche) %>% 
  adorn_percentages("row") %>% 
  adorn_pct_formatting(digits = 1)

```

FI_orNot Variable
```{r}





#A table 
data_dec21 %>% filter(!is.na(FI_orNot)) %>% 
tabyl(Sex, FI_orNot) %>% 
  adorn_percentages("row") %>% 
  adorn_pct_formatting(digits = 1)



#subset(data_dec21, Race == "White Only") 



#data_dec21 %>% filter(!is.na(FI_orNot), FI_orNot == #"Food Insecure") %>% 
#  ggplot(aes(x = Sex, y = )) +
#  geom_bar(aes(y = ..prop.., fill = factor(..x..)),stat #= count) +
#  scale_y_continuous(label = scales::percent)
  


```








## Merging Data sets

Importing 2nd data set

```{r}
#Read in xlsx file
poverty_data_large <- read_excel("C:/Users/foste/Desktop/418 final project folder/PovertyEstimates_new.xlsx")

poverty_data_county <- select(poverty_data_large,FIPS_code, PCTPOVALL_2020)
#rm(poverty_data_large)

```

Merging Data sets by inner join

```{r}
Merged_DataSet <-  data_dec21 %>% inner_join(poverty_data_county,by="FIPS_code")
rm(poverty_data_large)

```


Merging these data sets will tell us for each observation, what the percentage of people in the county live in poverty

```{r}


```















## New Data Set, % in poverty by State

```{r, warning = FALSE}
#Pct_in_poverty_State <- read.csv("C:/Users/foste/Desktop/418 final project folder/Pct_in_poverty_State - Sheet1.csv")

#summary(Pct_in_poverty_State)
```

Pivot Longer to change shape of Data Set

```{r}
#Pct_in_poverty_State <- Pct_in_poverty_State %>%
#pivot_longer(names_from = Stabr, values_from = Pct_in_poverty)
```


```{r}

#Pct_in_poverty_State %>% ggplot(aes(x = Stabr, y = pct_in_poverty)) +
#geom_col(fill = "dodgerblue", color = "black")+
#labs(title = "% in poverty by State",
#caption = "The American Community servey",
#x = "State",
#y = "% in Poverty")


```

